---
title: "Getting started with houdini"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with houdini}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

houdini injects SAS-style RTF tables and figures into a Word (`.docx`) template
at named bookmarks. The primary interface is a Shiny app, but the underlying
pipeline is a set of plain R functions. This vignette walks through both: first
the interactive workflow, then the pipeline the app is built on, using a small
self-contained RTF so the examples run anywhere.

## The interactive workflow

Launch the app with:

```{r, eval = FALSE}
library(houdini)
run_app()
```

The app has three inputs on the left:

1. **Word Document** — a `.docx` template containing named bookmarks. houdini
   reports how many bookmarks it finds. Bookmarks whose names begin with `_`
   (Word's internal bookmarks) are ignored.
2. **RTF Source** — a folder of `.rtf` files, one per table or figure. The file
   name (without extension) becomes the *table name* you map bookmarks to.
3. **Generate** — produces the finished document and a generation log.

In the middle, the **Table Configuration** grid maps each bookmark to a table.
Clicking a row loads that table into the preview on the right, where you can
filter by parameter/timeline and click columns or rows to exclude them.

For large jobs, define the mapping in a spreadsheet and use **Import Excel**
instead of filling the grid by hand — see `?houdini_app` and the README for the
column format.

## The pipeline, step by step

The app is a thin layer over a handful of pipeline functions. To see them work
without needing a real Word file, we build a tiny two-page RTF by hand. This is
the same shape SAS ODS emits: a repeating header row, a `Parameter:` line in the
page header, and timeline-label rows (`Week 1`, `Week 4`) separating blocks of
data.

```{r}
rtf <- r"(
{\rtf1\ansi
\sectd
{\header \trowd\cellx9999{Parameter: Cholesterol\cell}\row}
\trowd\trhdr\cellx1440\cellx2520\cellx3600
{\b Treatment\cell}{\b N\cell}{\b Mean\cell}\row
\trowd\cellx1440\cellx2520\cellx3600
{Week 1\cell}{\cell}{\cell}\row
\trowd\cellx1440\cellx2520\cellx3600
{Placebo\cell}{50\cell}{5.2\cell}\row
\trowd\cellx1440\cellx2520\cellx3600
{Drug A\cell}{48\cell}{4.8\cell}\row
\sect
\sectd
{\header \trowd\cellx9999{Parameter: Triglycerides\cell}\row}
\trowd\trhdr\cellx1440\cellx2520\cellx3600
{\b Treatment\cell}{\b N\cell}{\b Mean\cell}\row
\trowd\cellx1440\cellx2520\cellx3600
{Week 4\cell}{\cell}{\cell}\row
\trowd\cellx1440\cellx2520\cellx3600
{Placebo\cell}{49\cell}{5.5\cell}\row
\sect
}
)"

path <- tempfile(fileext = ".rtf")
writeLines(rtf, path)
```

### Inspecting a table

`houdini:::get_table_info()` summarises a table the way the app does when you
click a configuration row — its shape, and the parameter and timeline values
available to filter on:

```{r}
info <- houdini:::get_table_info(path)
str(info)
```

The two `Parameter:` header lines surface as filterable parameters, and the
`Week 1` / `Week 4` label rows surface as timelines.

### Rendering a preview

The preview panes are HTML. `get_table_html()` renders the whole table; the
filtered variant applies exclusions and filters exactly as the download will:

```{r}
html <- houdini:::get_table_html(path)
substr(html, 1, 200)
```

```{r}
# Keep only the Cholesterol page, drop the third column
html_filtered <- houdini:::get_table_html_output(
  path,
  parameters    = "Cholesterol",
  excluded_cols = 3
)
grepl("Triglycerides", html_filtered)  # dropped
```

### Producing the Word table XML

`get_table_xml()` is what actually goes into the document — a self-contained
`<w:tbl>` OOXML fragment:

```{r}
xml <- houdini:::get_table_xml(path, parameters = "Cholesterol")
# It parses as valid XML
!is.null(xml2::read_xml(xml))
```

### Injecting into a document

The final step needs a real `.docx` template with a matching bookmark, so it is
not run here, but the shape is:

```{r, eval = FALSE}
session <- houdini:::open_docx("template.docx")
houdini:::inject_table(session, "MyBookmark", xml)
houdini:::close_docx(session, "output.docx")
```

`process_document()` wraps all of this — parse, filter, render, inject — over a
whole configuration table, which is exactly what the **Generate** button calls.

## Where to go next

- `?run_app` / `?houdini_app` — the app entry points.
- The README — installation and the Excel configuration format.
